## Snowy Plover Cycles, Script
# DF for Temperature Maximum
2006_2016_T_max : All Temperature Max. values 2006-2016.
contains:
Y Year
D Date
TMAX Maximum Temp.
```{r}
tmax <- read.csv("/Users/Silvia/SnowyPlover/2006_2016_T_max.csv", header = TRUE, sep = ";", dec = ",")
View(tmax)
tmax$D <- as.Date(tmax$D, "%Y-%m-%d")
```
#DF for Days after Springtide:
contains:
TH Tideheight in cm
ST Event of Springtide, with Springtide-Height. Always at full moon and new moon.
DaST Days after Springtide
MP Time of Meridian Passing (noon of moon)
IMP Illumination of Meridian Passing
NM/FM Time of Newmoon and Fullmoon, correlates with spring tide
DST Date of Springtide, else NA
```{r}
st <- read.csv("/Users/Silvia/SnowyPlover/Days_after_Springtide_DST.csv", header = TRUE, sep = ";", dec = ",")
View(st)
st$D <- as.Date(st$D, "%Y-%m-%d")
st$DST <- as.Date(st$DST, "%Y-%m-%d")
```
# DF for interpolated Illumination at 12 pm (iin)
contains:
Date
Inoon interpolated Illumination at 12 pm for given date
```{r}
iin <- read.csv("/Users/Silvia/SnowyPlover/Illumination_Noon.csv", header = TRUE, sep = ";", dec = ",")
View(iin)
iin$Date <- as.Date(iin$Date, "%Y-%m-%d")
#install.packages("formattable")
library(formattable)
iin$Inoon <- percent(iin$Inoon)
```
# DF for Days after Newmoon
contains:
Year
Date
MP Time of Meridian Passing (noon of moon)
IMP Illumination at Meridian Passing
DaNM Days after last Newmoon
msequ Phase of the Moon (NM: New Moon, WxC: Waxing Crescent, FQ: First Quater, WxG: Waxing Gibbous, FM: Full Moon, WnG: Waning Gibbous, TQ: Third Quater, WnC: Waning Crescent)
```{r}
nm <- read.csv("/Users/Silvia/SnowyPlover/Days_after_NM_msequ.csv", header = TRUE, sep = ";", dec = ",")
View(nm)
nm$Date <- as.Date(nm$Date, "%Y-%m-%d")
library(chron)
nm$MP <- chron(times=nm$MP) # telling R that that's a time format
nm$NM <- chron(times= nm$NM)
library(formattable)
nm$IMP <- percent(nm$IMP) # change the Illumination in percentage
```
# New Hatch Data 2006-2016 (excl.2014), nhd
contains:
Year
ID Bird ID
FD Found Date
LD Laying Date
ED End Date
F Fate
```{r}
nhd <- read.csv("/Users/Silvia/SnowyPlover/Hatchdata_2006_2016.csv", header = TRUE, sep = ";", dec = ",")
View(nhd)
nhd$Site <- NULL
nhd$Nest <- NULL
nhd$FD <- as.Date(nhd$FD, "%Y-%m-%d") # Founddate
nhd$LD <- as.Date(nhd$LD, "%Y-%m-%d") # Laying Date
nhd$ED <- as.Date(nhd$ED, "%Y-%m-%d") # Enddate
```
merge the dfs together:
```{r}
nhd$T <- tmax$TMAX[match(nhd$LD,tmax$D)] # Maximum Temp. at Laying Date
nhd$MP <- nm$MP[match(nhd$LD,nm$Date)] # Meridian Passing at Laying Date, NAs indicate, that there was no MP at this day.
nhd$DaNM <- nm$DaNM[match(nhd$LD,nm$Date)] # Days after last Newmoon at Laying Date
nhd$Inoon <- iin$Inoon[match(nhd$LD,iin$Date)] # Illumination at 12pm at Laying Date
nhd$TH <- st$TH[match(nhd$LD,st$D)] # Tideheight in cm at Laying Date
nhd$ST <- st$ST[match(nhd$LD,st$D)] # Event of Springtide, in Height of Springtide at Laying Date
nhd$DaST <- st$DaST[match(nhd$LD,st$D)] # Days after Springtide
nhd$IMP <- nm$IMP[match(nhd$LD,nm$Date)] # Illumination at Merdian Passing
nhd$msequ <- nm$msequ [match(nhd$LD,nm$Date)] # Phase of the Moon at Laying Date
nhd$DST <- st$DST [match(nhd$LD,st$D)] # Date of Springtide if there was a Springtide
```
Standardize the Dates:
transform in Julian Days, centered by mean and divided by sd for each year
MESSY WORKAROUND
```{r}
#The function for that is scale(), but as I don`t want the extra rows with attributes I hashtagged the commandos for it in the source code and created a new function scaleNoAttr:

scaleNoAttr <- function (x, center = TRUE, scale = TRUE) 
{ 
    x <- as.matrix(x) 
    nc <- ncol(x) 
    if (is.logical(center)) { 
        if (center) { 
            center <- colMeans(x, na.rm = TRUE) 
            x <- sweep(x, 2L, center, check.margin = FALSE) 
        } 
    } 
    else if (is.numeric(center) && (length(center) == nc)) 
        x <- sweep(x, 2L, center, check.margin = FALSE) 
    else stop("length of 'center' must equal the number of columns of 'x'") 
    if (is.logical(scale)) { 
        if (scale) { 
            f <- function(v) { 
                v <- v[!is.na(v)] 
                sqrt(sum(v^2)/max(1, length(v) - 1L)) 
            } 
            scale <- apply(x, 2L, f) 
            x <- sweep(x, 2L, scale, "/", check.margin = FALSE) 
        } 
    } 
    else if (is.numeric(scale) && length(scale) == nc) 
        x <- sweep(x, 2L, scale, "/", check.margin = FALSE) 
    else stop("length of 'scale' must equal the number of columns of 'x'") 
    #if (is.numeric(center)) 
    #    attr(x, "scaled:center") <- center 
    #if (is.numeric(scale)) 
    #    attr(x, "scaled:scale") <- scale 
    x 
} 
scale1<- function (x, center = TRUE, scale = TRUE) 
{ 
    x <- as.matrix(x) 
    nc <- ncol(x) 
    if (is.logical(center)) { 
        if (center) { 
            center <- colMeans(x, na.rm = TRUE) 
            x <- sweep(x, 2L, center, check.margin = FALSE) 
        } 
    } 
    else if (is.numeric(center) && (length(center) == nc)) 
        x <- sweep(x, 2L, center, check.margin = FALSE) 
    else stop("length of 'center' must equal the number of columns of 'x'") 
    if (is.logical(scale)) { 
        if (scale) { 
            f <- function(v) { 
                v <- v[!is.na(v)] 
                sqrt(sum(v^2)/max(1, length(v) - 1L)) 
            } 
            scale <- apply(x, 2L, f) 
            x <- sweep(x, 2L, scale, "/", check.margin = FALSE) 
        } 
    } 
    else if (is.numeric(scale) && length(scale) == nc) 
        x <- sweep(x, 2L, scale, "/", check.margin = FALSE) 
    else stop("length of 'scale' must equal the number of columns of 'x'") 
    #if (is.numeric(center)) 
    #    attr(x, "scaled:center") <- center 
    #if (is.numeric(scale)) 
    #    attr(x, "scaled:scale") <- scale 
    x 
} 

# transform dates in julian days

require(lubridate)
nhd$EDY <- yday(nhd$ED) # Enddate
nhd$FDY <- yday(nhd$FD) # Founddate
nhd$LDY <- yday(nhd$LD) # Laying Date
nhd$DSTY <- yday(nhd$DST) # Date of Springtide

# scaling of the Julian Dates. ("z" for normalization)
m <- split(nhd$EDY, nhd$Year)  # splitting the data into years
m6 <- scaleNoAttr(m$`2006`) # creating vectors with standardized dates for every year
m7 <- scaleNoAttr(m$`2007`)
m8 <- scaleNoAttr(m$`2008`)
m9 <- scaleNoAttr(m$`2009`)
m10 <- scaleNoAttr(m$`2010`)
m11 <- scaleNoAttr(m$`2011`)
m12 <- scaleNoAttr(m$`2012`)
m13 <- scaleNoAttr(m$`2013`)
m15 <- scaleNoAttr(m$`2015`)
m16 <- scaleNoAttr(m$`2016`)

lst <- list(m6[,1],m7[,1],m8[,1],m9[,1],m10[,1],m11[,1],m12[,1],m13[,1], m15[,1], m16[,1] ) 
vec <- Reduce(c,lst) # transform those vectors into one vector
nhd$EDZ <- NA # new col 
nhd$EDZ <- vec # working! But not very elegant.


b <- split(nhd$FDY, nhd$Year)  # splitting the data into years
b6 <- scaleNoAttr(b$`2006`) # creating vectors with standardized dates for every year
b7 <- scaleNoAttr(b$`2007`)
b8 <- scaleNoAttr(b$`2008`)
b9 <- scaleNoAttr(b$`2009`)
b10 <- scaleNoAttr(b$`2010`)
b11 <- scaleNoAttr(b$`2011`)
b12 <- scaleNoAttr(b$`2012`)
b13 <- scaleNoAttr(b$`2013`)
b15 <- scaleNoAttr(b$`2015`)
b16 <- scaleNoAttr(b$`2016`)

lst <- list(b6[,1],b7[,1],b8[,1],b9[,1],b10[,1],b11[,1],b12[,1],b13[,1], b15[,1], b16[,1] ) 
vec <- Reduce(c,lst) # transform those vectors into one vector
nhd$FDZ <- NA # new col 
nhd$FDZ <- vec 


g <- split(nhd$LDY, nhd$Year)  # splitting the data into years
g6 <- scaleNoAttr(g$`2006`) # creating vectors with standardized dates for every year
g7 <- scaleNoAttr(g$`2007`)
g8 <- scaleNoAttr(g$`2008`)
g9 <- scaleNoAttr(g$`2009`)
g10 <- scaleNoAttr(g$`2010`)
g11 <- scaleNoAttr(g$`2011`)
g12 <- scaleNoAttr(g$`2012`)
g13 <- scaleNoAttr(g$`2013`)
g15 <- scaleNoAttr(g$`2015`)
g16 <- scaleNoAttr(g$`2016`)

lst <- list(g6[,1],g7[,1],g8[,1],g9[,1],g10[,1],g11[,1],g12[,1],g13[,1], g15[,1], g16[,1] ) 
vec <- Reduce(c,lst) # transform those vectors into one vector
nhd$LDZ <- NA # new col 
nhd$LDZ <- vec 


p <- split(nhd$DSTY, nhd$Year)  # splitting the data into years
p6 <- scaleNoAttr(p$`2006`) # creating vectors with standardized dates for every year
p7 <- scaleNoAttr(p$`2007`)
p8 <- scaleNoAttr(p$`2008`)
p9 <- scaleNoAttr(p$`2009`)
p10 <- scaleNoAttr(p$`2010`)
p11 <- scaleNoAttr(p$`2011`)
p12 <- scaleNoAttr(p$`2012`)
p13 <- scaleNoAttr(p$`2013`)
p15 <- scaleNoAttr(p$`2015`)
p16 <- scaleNoAttr(p$`2016`)

lst <- list(p6[,1],p7[,1],p8[,1],p9[,1],p10[,1],p11[,1],p12[,1],p13[,1], p15[,1], p16[,1] ) 
vec <- Reduce(c,lst) # transform those vectors into one vector
nhd$DSTz <- NA # new col 
nhd$DSTz <- vec 
```



